---
title: "4-November-2024"
format: html
editor: visual
---

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

import the data
```{r}
forest <- read.csv(here("data", "Forest_Tree_DBH_F21.csv"))

forest$Forest <- as.factor(forest$Forest)
```

Question 1: Is there a difference in mean tree DBH between forests

Question 2: If there is a difference, for which forest is there a difference?


```{r}
ggplot(forest, aes(x=Forest, y = DBH_cm)) +
  geom_boxplot() +
  coord_flip() +
  geom_jitter(width=.1)
```

I do not think that there will be a difference between any 2 means

create a lm:
```{r}
forest_model <- lm(DBH_cm ~ Forest, data = forest)
```

check assumptions
```{r}
autoplot(forest_model, smooth.colour=NA)
```

run ANOVA test:
```{r}
anova(forest_model)
```

check summary table
```{r}
summary(forest_model)
```
We cant see which forests are different from other forests, only which forests are different than degrasse


```{r}
library(multcomp)
```

Conduct a tukeys HSD test (runs every forest against every other forest)
```{r}
phc1 <- glht(forest_model, linfct = mcp(Forest="Tukey"))
summary(phc1)
```

```{r}
cld(phc1) # compact letter display
# if groups share the same letter then they are not significantly different from each other
```

only south hammond and degrasse have a significant difference in mean DBH