---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---


```{r}
rm(list=ls())
library(tidyverse)
library(here)

bmi <- read.csv(here("data", "body_fat_percentage.csv"))
glimpse(bmi)

bmi$gender <- as.factor(bmi$gender)
```

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships
```{r}
ggplot(bmi, aes(body_fat_percent)) +
  geom_histogram(binwidth=4) +
  facet_wrap(bmi$gender, ncol = 1)
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.
```{r}
bmi_data <- bmi |> group_by(gender) |>
  summarise(mean = mean(body_fat_percent, na.rm = TRUE),
            sd = sd(body_fat_percent, na.rm = TRUE))
```

3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?

Equal variance in 2 groups, independence, and normalcy of residuals. This ensures that the results returned by the model are reliable. I think that they appear to meet this assumption

  - Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
  
  Yes, i expect there to be a significant difference between mean body fat percentage
    
    
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?
```{r}
t.test(body_fat_percent ~ gender, data = bmi)
# t.test(response ~ predictor, data = data frame)
```
Men have significantly less body fat than women (t = 2.89, df = 20.96, p = .00932)

df = sample size - response variables
